Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(ggplot2)
library(gridExtra)
summary(pf)
p1 <- ggplot(subset(pf, !is.na(gender)), aes(age, friend_count)) +
  geom_line(aes(color = gender))


p2 <- ggplot(subset(pf, !is.na(gender)), aes(age, friend_count)) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)

grid.arrange(p1, p2, ncol= 1)
```

# Write code to create a new data frame,
# called 'pf.fc_by_age_gender', that contains
# information on each age AND gender group.

# The data frame should contain the following variables:

#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)

# Here is an example of the structure of your data frame. Your
# data values will be different. Note that if you are grouping by
# more than one variable, you will probably need to call the
# ungroup() function. 

#   age gender mean_friend_count median_friend_count    n
# 1  13 female          247.2953                 150  207
# 2  13   male          184.2342                  61  265
# 3  14 female          329.1938                 245  834
# 4  14   male          157.1204                  88 1201

# See the Instructor Note for two hints.

# DO NOT DELETE THESE NEXT TWO LINES OF CODE
# ==============================================================
pf <- read.delim('/datasets/ud651/pseudo_facebook.tsv')
suppressMessages(library(dplyr))

# ENTER YOUR CODE BELOW THIS LINE.
# ==============================================================

***

``` {r}
library(dplyr)
pf.fc_by_age_gender <- summarise(group_by(filter(pf, !is.na(gender)), age, gender),
                                 mean_friend_count = mean(friend_count),
                                 median_friend_count = median(friend_count),
                                 n = n())
pf.fc_by_age_gender <- arrange(pf.fc_by_age_gender, age)
```

### Plotting Conditional Summaries
Notes:
# Create a line graph showing the
# median friend count over the ages
# for each gender. Be sure to use
# the data frame you just created,
# pf.fc_by_age_gender.

# See the Instructor Notes for a hint.

# This assignment is not graded and
# will be marked as correct when you submit.

# ENTER YOUR CODE BELOW THIS LINE
# =================================================

```{r Plotting Conditional Summaries}
ggplot(pf.fc_by_age_gender, aes(age, median_friend_count)) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

***
# Plot the ratio of the female to male median
# friend counts using the data frame
# pf.fc_by_age_gender.wide.

# Think about what geom you should use.
# Add a horizontal line to the plot with
# a y intercept of 1, which will be the
# base line. Look up the documentation
# for geom_hline to do that. Use the parameter
# linetype in geom_hline to make the
# line dashed.

# The linetype parameter can take the values 0-6:
# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash

# This assignment is not graded and
# will be marked as correct when you submit.

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)
```

```{r}
library(tidyr)
library(dplyr)

pf.fc_by_age_gender.wide <- select(pf.fc_by_age_gender,
                                  c('age', 'gender', 'median_friend_count')) %>%
                            filter(!is.na(gender)) %>%
                            spread(key = gender, value = median_friend_count) %>%
                            mutate(ratio = female/male)
```

***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(pf.fc_by_age_gender.wide, aes(age, ratio)) +
  geom_line() +
  geom_hline(yintercept = 1, color = 'orange', linetype = 1, alpha = .6)
```

***

### Third Quantitative Variable
Notes:
# Create a variable called year_joined
# in the pf data frame using the variable
# tenure and 2014 as the reference year.

# The variable year joined should contain the year
# that a user joined facebook.

```{r Third Quantitative Variable}
pf.year_joined <- mutate(pf, year_joined = floor(2014 - tenure / 365))
summary(pf.year_joined$year_joined)
```

***

### Cut a Variable
Notes:
# Create a new variable in the data frame
# called year_joined.bucket by using
# the cut function on the variable year_joined.

# You need to create the following buckets for the
# new variable, year_joined.bucket

#        (2004, 2009]
#        (2009, 2011]
#        (2011, 2012]
#        (2012, 2014]

# Note that a parenthesis means exclude the year and a
# bracket means include the year.
```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf.year_joined$year_joined, breaks = c(2004, 2009,2011,2012, 2014))
```

***

### Plotting it All Together
Notes:
# Create a line graph of friend_count vs. age
# so that each year_joined.bucket is a line
# tracking the median user friend_count across
# age. This means you should have four different
# lines on your plot.

# You should subset the data to exclude the users
# whose year_joined.bucket is NA.

```{r Plotting it All Together}
ggplot(subset(pf, !is.na(year_joined.bucket)), aes(age, friend_count)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:
# Write code to do the following:

# (1) Add another geom_line to code below
# to plot the grand mean of the friend count vs age.

# (2) Exclude any users whose year_joined.bucket is NA.

# (3) Use a different line type for the grand mean.

# As a reminder, the parameter linetype can take the values 0-6:

# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash
```{r Plot the Grand Mean}
ggplot(subset(pf, !is.na(year_joined.bucket)), aes(age, friend_count)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, pf$tenure >0), summary(friend_count / tenure))

```

***

### Friendships Initiated
Notes:
# Create a line graph of mean of friendships_initiated per day (of tenure)
# vs. tenure colored by year_joined.bucket.

# You need to make use of the variables tenure,
# friendships_initiated, and year_joined.bucket.

# You also need to subset the data to only consider user with at least
# one day of tenure.

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
select(pf, friendships_initiated, year_joined.bucket, tenure) %>%
  filter(tenure > 1) %>%
  ggplot(aes(tenure, friendships_initiated/tenure)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket))
```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
summary(yo)

str(yo)
yo$id <- factor(yo$id)
str(yo)
length(unique(yo$price))

summary(yo$price)
ggplot(yo, aes(price)) +
  geom_histogram(binwidth = 1)
```


***

### Number of Purchases
Notes:
# Create a new variable called all.purchases,
# which gives the total counts of yogurt for
# each observation or household.

# One way to do this is using the transform
# function. You can look up the function transform
# and run the examples of code at the bottom of the
# documentation to figure out what it does.

# The transform function produces a data frame
# so if you use it then save the result to 'yo'!

# OR you can figure out another way to create the
# variable.

```{r Number of Purchases}
yo <- transform(yo, all.purchases = yo$strawberry + yo$blueberry + yo$pina.colada + yo$plain + yo$mixed.berry)
```

***

### Prices over Time
Notes:
# Create a scatterplot of price vs time.

# This will be an example of a time series plot.

# Resolve overplotting issues by using
# techniques you learned in Lesson 4.
```{r Prices over Time}
ggplot(yo, aes(yo$time, yo$price)) +
  geom_point(alpha=.3, fill=I('#ff0000'))
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(40000)
sample.ids <- sample(levels(yo$id), 16)
ggplot(subset(yo, id %in% sample.ids),
       aes(time, price)) +
  geom_line() +
  facet_wrap(~id) +
  geom_point(aes(size = all.purchases), pch = 1)
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

