Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***
```{r}
pf = read.delim('pseudo_facebook.tsv')
```
  

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
summary(pf$age)
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(pf, aes(age, friend_count)) +
  geom_point() +
  xlim(13, 90)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(pf, aes(age, friend_count)) +
  geom_point(alpha = .05) +
  xlim(13, 90)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(pf, aes(age, friend_count)) +
  geom_point(alpha = .05) +
  xlim(13, 90) +
  coord_trans(y = "sqrt")
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(pf, aes(age, friend_count)) +
  geom_point(alpha = .05) +
  xlim(13, 90) +
  coord_trans(y = "sqrt")
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(pf, aes(age, friendships_initiated)) +
  geom_point(alpha = .05, position = position_jitter(height = 0)) +
  coord_trans(y = "sqrt")
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)

pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(fcMean = mean(friend_count),
            fcMedian = median(friend_count),
            n = n()) %>%
  arrange(age)

head(pf.fc_by_age)
```

Create your plot!

```{r Conditional Means Plot}
ggplot(pf.fc_by_age, aes(age, fcMedian)) +
  geom_line(na.rm = TRUE) +
  geom_smooth() +
  scale_size_area()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(pf, aes(age, friend_count)) +
  geom_point(alpha = .05,
             position = position_jitter(h=0),
             color = 'orange') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = c(.1),
            linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.9),
            linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = c(.5), color = 'blue') +
  coord_cartesian(xlim = c(13, 75), ylim = c(0, 1000))
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count)
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:
# Create a scatterplot of likes_received (y)
# vs. www_likes_received (x). Use any of the
# techniques that you've learned so far to
# modify the plot.

```{r}
summary(pf$www_likes_received)
summary(pf$likes_received)

ggplot(pf, aes(likes_received, www_likes_received)) +
  geom_point(alpha = .05,
             position = position_jitter(h=0),
             color = 'orange') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = c(.1),
            linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.9),
            linetype = 2, color = 'blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = c(.5), color = 'blue') +
  coord_cartesian(xlim = c(0, quantile(pf$likes_received, .95)),
                  ylim = c(0, quantile(pf$www_likes_received, .95))) + 
  geom_smooth()
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
with(pf, cor.test(likes_received, www_likes_received))
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
mc <- data(Mitchell)
```

Create your plot!

```{r Temp vs Month}
summary(Mitchell)

ggplot(Mitchell, aes(Month, Temp)) +
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
with(Mitchell, cor.test(Month, Temp))
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

range(Mitchell$Month)


ggplot(Mitchell, aes(Month, Temp)) +
  geom_point() +
  scale_x_continuous(breaks = seq(0, max(Mitchell$Month), 12))
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:
```{r}
ggplot(Mitchell, aes(Month%%12, Temp)) +
  geom_point() +
  scale_x_continuous(breaks = seq(0, max(Mitchell$Month), 12))
```

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
p1 <- ggplot(subset(pf.fc_by_age, age < 71), aes(age, fcMedian)) +
  geom_line()
```

***

### Age with Months Means
# Create a new variable, 'age_with_months', in the 'pf' data frame.
# Be sure to save the variable in the data frame rather than creating
# a separate, stand-alone variable. You will need to use the variables
# 'age' and 'dob_month' to create the variable 'age_with_months'.

# Assume the reference date for calculating age is December 31, 2013.

# This programming assignment WILL BE automatically graded. For
# this exercise, you need only create the 'age_with_months' variable;
# no further processing of the data frame is necessary.

```{r Age with Months Means}
head(pf)
pf$age_with_months <- with(pf, age + (1 - dob_month / 12))
```

# Create a new data frame called
# pf.fc_by_age_months that contains
# the mean friend count, the median friend
# count, and the number of users in each
# group of age_with_months. The rows of the
# data framed should be arranged in increasing
# order by the age_with_months variable.

# For example, the first two rows of the resulting
# data frame would look something like...

# age_with_months  friend_count_mean	friend_count_median	n
#              13            275.0000                   275 2
#        13.25000            133.2000                   101 11

Programming Assignment
```{r Programming Assignment}
pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age_with_months)

head(pf.fc_by_age_months)

age_with_months_group <- group_by(pf, age_with_months)
pf.fc_by_age_months <- summarise(age_with_months_group,
                                 friend_count_mean = mean(friend_count),
                                 friend_count_median = median(friend_count),
                                 n = n())
pf.fc_by_age_months <- arrange(pf.fc_by_age_months, age_with_months)
head(pf.fc_by_age_months)
```

***

### Noise in Conditional Means
# Create a new line plot showing friend_count_mean versus the new variable,
# age_with_months. Be sure to use the correct data frame (the one you created
# in the last exercise) AND subset the data to investigate users with ages less
# than 71.
```{r Noise in Conditional Means}
p2 <- ggplot(subset(pf.fc_by_age_months, age_with_months < 71),
       aes(age_with_months, friend_count_mean)) +
  geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
library(gridExtra)
grid.arrange(p1, p2, ncol = 1)
```
```{r}
p1 <- p1 + geom_smooth()
p2 <- p2 + geom_smooth()
grid.arrange(p1, p2, ncol = 1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

